---
title: "Data Profiling"
runtime: shiny
output: html_document   
--- 


```{r echo = FALSE}
#rmarkdown::run('src/shiny-markdown/data-profiling-interactive.Rmd')
#profileFor2010 <- "/sfs/qumulo/qhome/itm3f/git/dspg20STW/src/shiny-markdown/2010.csv"
#profileFor2011 <- "/sfs/qumulo/qhome/itm3f/git/dspg20STW/src/shiny-markdown/2011.csv"

#profileFor2010 <- "src/shiny-markdown/2010.csv"
#profileFor2011 <- "src/shiny-markdown/2011.csv"
#profileFor2012 <- "src/shiny-markdown/2012.csv" 
#profileFor2013 <- "src/shiny-markdown/2013.csv" 

prof2010 <- "data/ncses_stw/profiling_tables/prof2010.csv"
prof2011 <- "data/ncses_stw/profiling_tables/prof2011.csv"
prof2012 <- "data/ncses_stw/profiling_tables/prof2012.csv"
prof2013 <- "data/ncses_stw/profiling_tables/prof2013.csv"
prof2014 <- "data/ncses_stw/profiling_tables/prof2014.csv"
prof2015 <- "data/ncses_stw/profiling_tables/prof2015.csv"
prof2016 <- "data/ncses_stw/profiling_tables/prof2016.csv"
prof2017 <- "data/ncses_stw/profiling_tables/prof2017.csv"
prof2018 <- "data/ncses_stw/profiling_tables/prof2018.csv"
prof2019 <- "data/ncses_stw/profiling_tables/prof2019.csv"





selectInput("profiles", label = "Profile", choices = c(2010,2011,2012,2013,2014,2015,2016,2017,2018,2019))
 
```

```{r echo = FALSE} 

myCSV <- reactive({
  if(input$profiles == 2010){
    read.csv(prof2010)
  }else if(input$profiles == 2011){ 
    read.csv(prof2011)
  }else if(input$profiles == 2012){
    read.csv(prof2012)  
  }else if(input$profiles == 2013){
    read.csv(prof2013)
  }else if(input$profiles == 2014){
    read.csv(prof2014)
  }else if(input$profiles == 2015){
    read.csv(prof2015)
  }else if(input$profiles == 2016){
    read.csv(prof2016)
  }else if(input$profiles == 2017){
    read.csv(prof2017)
  }else if(input$profiles == 2018){
    read.csv(prof2018)
  }else{
    read.csv(prof2019)
  }
  })

renderTable({
  
  myCSV()
  
  
}) 


```

###Variables
*note: All information regarding variable descriptions was taken from the Burning Glass Data Dictionary*

* id - unique ID generated by BGT 
* jobdate - date the posting was spidered 
* state - state where job is located
* soc - code of the job assigned; these are always the first 6 digits of a job's O*NET code
* socname - name for a posting's SOC code
* lat - latitude
* lon - longitude
* minedu - minimum education requirement
* maxedu - maximum education  

###Metrics

* Completeness: The percentage of how complete the data is
* Validity: The percentage of data elements whose attributes possess values within the range expected for a legitimate entry
* Uniqueness: number of unique valid values that have been entered for a variable
