---
title: "Data Profiling"
runtime: shiny
output: html_document
---


```{r echo = FALSE}
#rmarkdown::run('src/shiny-markdown/data-profiling-interactive.Rmd')
#profileFor2010 <- "/sfs/qumulo/qhome/itm3f/git/dspg20STW/src/shiny-markdown/2010.csv"
#profileFor2011 <- "/sfs/qumulo/qhome/itm3f/git/dspg20STW/src/shiny-markdown/2011.csv"

profileFor2010 <- "src/shiny-markdown/2010.csv"
profileFor2011 <- "src/shiny-markdown/2011.csv"
profileFor2012 <- "src/shiny-markdown/2012.csv"

selectInput("profiles", label = "Profile", choices = c(2010,2011,2012))

```

```{r echo = FALSE}

myCSV <- reactive({
  if(input$profiles == 2010){
    read.csv(profileFor2010)
  }else if(input$profiles == 2011){
    read.csv(profileFor2011)
  }else{
    read.csv(profileFor2012)
  }
  })

renderTable({
  
  myCSV()
  
  
}) 


```

###Variables
*note: All information regarding variable descriptions was taken from the Burning Glass Data Dictionary*

* id - unique ID generated by BGT
* jobdate - date the posting was spidered
* state - state where job is located
* soc - code of the job assigned; these are always the first 6 digits of a job's O*NET code
* socname - name for a posting's SOC code
* lat - latitude
* lon - longitude
* minedu - minimum education requirement
* maxedu - maximum education  


